---
title: "STAT504_Q2"
author: "Houjie Wang & Shuang Wu"
date: "2/19/2021"
output: pdf_document
---

```{r, include=FALSE}
library(nnet)
library(Rfast)
```

# Problem Recap

Are more physical activities related to a healthier life? 

# Related Variables Recap 

## Variables 

HEALTH: If -9, not ascertained; if -8, I don’t know; if -7, prefer not to answer; if 1, excellent; if 2, very good; if 3, good; if 4, fair; if 5, poor

## Predictor 

\begin{itemize}
  \item[LPACT] Count of times of light or moderate physical activity in past week
  \item[VPACT] Count of times of vigorous physical activity in past week
  \item[PHYACT] If -9, not ascertained; if -8, I don’t know; if -7, prefer not to answer; if 1, rarely or never conduct any physical activities; if 2, some light or moderate physical activities; 3. Some vigorous physical activities
\end{itemize} 

# Data Cleaning 

To simplify, observations only with positive values in HEALTH and PHYCT are taken. 
```{r, message=FALSE}
data <- read.csv("DataforUse.csv")
# data_full <- read.csv("perpub.csv")
valid_health_ind <- data$HEALTH %in% 1: 5 + data$PHYACT %in% 1: 3 + 
  as.vector(data$LPACT >= -1) + as.vector(data$VPACT >= -1) == 4
data_health <- cbind.data.frame(HEALTH = data$HEALTH[valid_health_ind], 
                                LPACT = as.integer(data$LPACT[valid_health_ind]), 
                                VPACT = as.integer(data$VPACT[valid_health_ind]), 
                                PHYACT = as.factor(data$PHYACT[valid_health_ind]))
nrow(data_health)

# Verify if LPACT=VPACT=-1, when PHYACT=1 (TRUE)
sum(rowsums(as.matrix(data_health[data_health$PHYACT == 1, 2: 3])) != -2) == 0

# Verify if VPACT=-1, when PHYACT=2 (TRUE)
sum(data_health[data_health$PHYACT == 2, 3] != -1) == 0

# Verify if LPACT=-1, when PHYACT=3 (TRUE)
sum(data_health[data_health$PHYACT == 3, 2] != -1) == 0

# Set all "-1" in LPACT and VPACT to "0" since they both represent no LPACT or VPACT last week
data_health$LPACT[data_health$LPACT == -1] = 0
data_health$VPACT[data_health$VPACT == -1] = 0
```  

# Model Proposal 
$$
\begin{aligned}
  HEALTH = \beta_0+\beta_1 LPACT + \beta_2VPACT
\end{aligned}
$$

Here is our multinomial logistic regression model. $Y$ represents HEALTH and $X$ represents the design matrix.
$$
\begin{aligned}
  P(Y_i=1)&=1-\sum_{k=2}^5 P(Y_i=k)=\frac{1}{1+\sum_{k=2}^5 \exp\{\boldsymbol{\beta}_k'\boldsymbol{X}_{i}\}}\\
  P(Y_i=2)&=P(Y_i=1)\exp\{\boldsymbol{\beta}_2'\boldsymbol{X}_{i}\}=\frac{\exp\{\boldsymbol{\beta}_2'\boldsymbol{X}_{i}\}}{1+\sum_{k=2}^5 \exp\{\boldsymbol{\beta}_k'\boldsymbol{X}_{i}\}} \\
  P(Y_i=3)&=P(Y_i=1)\exp\{\boldsymbol{\beta}_3'\boldsymbol{X}_{i}\}=\frac{\exp\{\boldsymbol{\beta}_3'\boldsymbol{X}_{i}\}}{1+\sum_{k=2}^5 \exp\{\boldsymbol{\beta}_k'\boldsymbol{X}_{i}\}} \\
  \vdots\\
  P(Y_i=5)&=P(Y_i=1)\exp\{\boldsymbol{\beta}_5'\boldsymbol{X}_{i}\}=\frac{\exp\{\boldsymbol{\beta}_5'\boldsymbol{X}_{i}\}}{1+\sum_{k=2}^5 \exp\{\boldsymbol{\beta}_k'\boldsymbol{X}_{i}\}}.
\end{aligned}
$$  

So we obtiant he optimization problem:
$$
\begin{aligned}
  \min_{\boldsymbol{\beta_2},\ldots,\boldsymbol{\beta_5}} \sum_{i=1}^n\sum_{k=1}^5\log P(Y_i=k)\boldsymbol{1}_{\{Y_i=k\}}.
\end{aligned}
$$

Now we fit such a model:
```{r}
fit <- multinom(HEALTH ~ LPACT + VPACT, data = data_health)
(results <- summary(fit))
``` 

After the model fit, we would like to evaluate the significance of coefficients. Here we use Wald test.  
```{r}
(1 - pnorm(abs(results$coefficients/results$standard.errors), 0, 1)) * 2
``` 

All p-values are extremely small and we conclude that the model is significant. Instead of evaluating coefficient significance, we are also interested in testing the robustness of the model as a whole. Here we use likelihood ratio test by comparing the proposed model to an empty model: 

```{r}
fit2 <- multinom(HEALTH ~ 1, data = data_health)
chi_stat <- -fit$deviance - (-fit2$deviance)
1 - pchisq(chi_stat, df = 2)
``` 

As we can see that the p-value is also small. The proposed model is strong.

<!-- fit2 <- multinom(HEALTH ~ LPACT + VPACT + LPACT*PHYACT + VPACT*PHYACT, data = data_health) -->
<!-- results2 <- summary(fit2) -->
<!-- (1 - pnorm(abs(results2$coefficients/results2$standard.errors), 0, 1)) * 2 -->
<!-- ```  -->

<!-- # Variable Selection  -->

<!-- ```{r} -->
<!-- library(parallel) -->
<!-- ff <- HEALTH ~ LPACT + VPACT + PHYACT + PHYACT*LPACT + PHYACT*VPACT -->
<!-- sub_ind <- sample(1: nrow(data_health), size = nrow(data_health), replace = FALSE) -->
<!-- data_health_sub <- data_health[sub_ind, ] -->
<!-- X <- model.matrix(ff, data_health_sub) -->
<!-- Y <- data_health_sub$HEALTH -->
<!-- fit_lasso_no <- glmnet(X, Y, family = "multinomial", alpha = 1, nlambda = 1, lambda = c(0), type.multinomial = "grouped") -->
<!-- coef(fit_lasso_no) -->

<!-- makeCluster(detectCores()) -->
<!-- start <- Sys.time() -->
<!-- fit_lasso <- glmnet::cv.glmnet(X, Y, family = "multinomial") -->
<!-- end <- Sys.time() -->
<!-- end - start -->
<!-- ```  -->





